### 2017-11-09 paper/sections/1_introduction.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the Mantle
language and policy engine to dynamically explore a variety of techniques,
ranging from basic algorithms and heuristics to statistical models, calculus,
and machine learning. While Mantle was originally designed for distributed file
systems, we show how the collection of abstractions effectively decomposes the
problem into manageable policies for a different domain and service (in this
case, cache management).  Our exploration of this space results in a two policy
scheme that achieves 96\% efficiency while using only 7.6\% of the memory
resources required by the base case.

@@

The fine-grained data annotation capabilities provided by key-value storage is
a natural match for many types of scientific simulation. Simulations relying on
a mesh-based decomposition of a physical region may result in millions or
billions of mesh cells. Each cell contains materials, pressures, temperatures
and other characteristics that are required to accurately simulate phenomena of
interest. In our target application, the
ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
hierarchy of cache nodes and a single node key-value store are used to store
both observed minima across a molecule's equation of motion (EOM) and the
hundreds or thousands of partial trajectories calculated each second during a
parallel job. Unfortunately, if we scale the system the IO to the storage
hierarchy will quickly saturate both the storage and bandwidth capacity of a
single node.

@@

In this paper we present a detailed analysis of how the ParSplice application
accesses key-value pairs over the course of a long running simulation across a
variety of initial conditions.  This type of analysis (1) shows the capacity
and resource requirements of a single node and (2) will help inform our load
balancing policies for when we switch to a distributed key-value store back-end
to store EOM minima.  For example, Figure~\ref{fig:motivation-regimes} shows
the cache activity of a ParSplice node and demonstrates that small changes to
the rates at which new atoms enter the simulation (\(\Delta\)) can have a
strong effect on the timing and frequency with which new EOM minima are
discovered and referenced.  The ParSplice caches on each node are trimmed when
memory pressure reaches a threshold but the number of unique keys accessed
(black points along bottom of both graphs) in
Figure~\ref{fig:motivation-regimes} suggests that a small cache of EOM minima
should be sufficient.

% What is Mantle
We use the Mantle language and policy engine to dynamically explore the effects
of different cache management strategies for the changing key-value workloads
generated by ParSplice.  Mantle~\cite{sevilla:sc15-mantle} has already proven
to be a critical control plane for improving file system metadata load
balancing  and in this work we show that the collection of abstractions
provided by Mantle is useful for reasoning about and designing different cache
management strategies.  Developers write policies for ``when" they want data
moved and ``how much" of the data to move, then the framework executes these
policies whenever a decision needs to be made.  These abstraction help
developers unfamiliar with the domain quickly reason about, develop, and deploy
new policies that control temporal and spatial locality. We show that Mantle:

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear keyspace access patterns and hot
spots that offer significant opportunity for optimization. We use the data
management and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service (in this case, cache management).  Our exploration
of this space results in a two policy scheme that achieves 96\% efficiency
while using only 7.6\% of the memory resources required by the base case.

@@

The fine-grained data annotation capabilities provided by key-value storage is
a natural match for many types of scientific simulation. Simulations relying on
a mesh-based decomposition of a physical region may result in millions or
billions of mesh cells. Each cell contains materials, pressures, temperatures
and other characteristics that are required to accurately simulate phenomena of
interest. In our target application, the
ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
hierarchy of cache nodes and a single node key-value store are used to store
both observed minima across a molecule's equation of motion (EOM) and the
hundreds or thousands of partial trajectories calculated each second during a
parallel job. Unfortunately, if we scale the system the IO to the storage
hierarchy will quickly saturate both the storage and bandwidth capacity of a
single node, so we need more effective data management techniques, such as
cache management or load balancing across a cluster.

@@

In this paper, we design cache management policies for the ParSplice worker
node, driven by a detailed analysis of the application's key-value accesses
over the course of a long running simulation across a variety of initial
conditions.  This type of analysis helps the system scale by (1) showing the
capacity and resource requirements of a single node and (2) informing the
design of load balancing policies for when we switch to a distributed key-value
store back-end to store EOM minima.  For example,
Figure~\ref{fig:motivation-regimes} shows the cache activity of a ParSplice
node and demonstrates that small changes to the rates at which new atoms enter
the simulation (\(\Delta\)) can have a strong effect on the timing and
frequency with which new EOM minima are discovered and referenced.  The
ParSplice caches on each node are trimmed when memory pressure reaches a
threshold but the number of unique keys accessed (black points along bottom of
both graphs) in Figure~\ref{fig:motivation-regimes} suggests that a small cache
of EOM minima should be sufficient.

% What is Mantle
We use the data management language and policy engine from the Mantle
paper~\cite{sevilla:sc15-mantle} to dynamically explore the effects of
different cache management strategies for the changing key-value workloads
generated by ParSplice.  Mantle was originally labeled a programmable file
system metadata load balancer, but we realize now that the collection of
abstractions was a control plane that improved metadata access. So in this
paper we refer to Mantle as a policy engine that injects policies written in
our data management language directly into a running service ({\it e.g.} file
system, key-value store, application).  We show how this approach is useful for
reasoning about and designing different cache management strategies.
Developers write policies for ``when" they want data moved and ``how much" of
the data to move, then the framework executes these policies whenever a
decision needs to be made.  These abstraction help developers unfamiliar with
the domain quickly reason about, develop, and deploy new policies that control
temporal and spatial locality. We show that Mantle:

### 2017-11-09 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
(MD) simulation package developed at LANL. It is part of the Exascale Computing
Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
Materials for the Future initiative. Its phases are:

@@

All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
cores per node.  Each node has 128GB of RAM and our goal is to limit the size
of the cache to 3\% of RAM\footnote{Empirically, we find this threshold works
well for most applications}. Note that this is an addition to the 30GB that
ParSplice uses to manage other ranks on the same node.  A single Cray node
produced trajectories that are \(5\times\) times longer than our 10 node
CloudLab clusters and \(25\times\) longer than UCSC's 10 node cluster. As a
result, it reaches different job phases faster and gives us a more
comprehensive view of the workload. The performance gains compared to the
commodity clusters have more to do with memory/PCI bandwidth than network.

@@

\subsubsection*{An active but small keyspace} The bars in
Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
worker tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when worker tasks generate segments for the same
state, which is computationally cheap (this motivates
Section~\S\ref{sec:static-load-balancing}).

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:the-need-for-dynamic-load-balancing-policies}).  The growth
rate, temperature, and number of workers have a predictable effect on the
structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different growth rates, but that the spatial
locality of key accesses is similar ({\it e.g.}, some keys are still
\(5\times\) more popular than others).  Figure~\ref{fig:motivation-regimes}
shows how entropy for different growth rates has temporal locality, as the
reads per second for \(\Delta_2\) looks like the reads per second for
\(\Delta_1\) stretched out along the time axis.  Trends also exist for
temperature and number of workers but are omitted here for space. This
structure means that we can learn the regimes and adapt the storage system
(this motivates Section~\S\ref{sec:ml-for-the-keyspace}).

>>>>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/parsplice.png}\\
\caption{The ParSplice architecture has a storage hierarchy of caches and a
dedicated cache node (purple boxes) backed by a persistent database (orange). A
splicer (S) tells worker nodes (W) to generate segments and workers employ
tasks (T) for more parallelization. We focus on the worker node's cache circled
in red, which facilitates communication and segment exchange between the worker
and its \label{fig:parsplice}} \end{figure}

ParSplice~\cite{perez:jctc20150parsplice} is an accelerated molecular dynamics
(MD) simulation package developed at LANL. It is part of the Exascale Computing
Project\footnote{http://www.exascale.org/bdec/} and is important to LANL's
Materials for the Future initiative. As shown in Figure~\ref{fig:parsplice},
the phases are:

@@

\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
and our goal is to limit the size of the cache to 3\% of
RAM\footnote{Empirically, we find this threshold works well for most
applications}. Note that this is an addition to the 30GB that ParSplice uses to
manage other ranks on the same node.  A single Cray node produced trajectories
that are \(5\times\) times longer than our 10 node CloudLab clusters and
\(25\times\) longer than UCSC's 10 node cluster. As a result, it reaches
different job phases faster and gives us a more comprehensive view of the
workload. The performance gains compared to the commodity clusters have more to
do with memory/PCI bandwidth than network.

The scalability experiment uses 1 splicer, persistent database node, and
in-memory database node and up to 2 workers. We scale up to 1024 tasks, which
spans 32 nodes and disable hyper-threading because we experience variability in
performance. For the rest of the experiments, we use 4 nodes, 1 splicer,
persistent database node, in-memory database node, worker, and up to 512 tasks.
The keyspace analysis is for the cache on the worker node, which is circled in
red in Figure~\ref{fig:parsplice}.

@@

\subsubsection*{An active but small keyspace} The bars in
Figure~\ref{fig:methodology-keyspace} show \(50-100\times\) as many reads
(\texttt{get()}) as writes (\texttt{put()}).  Worker tasks read the same key
for extended periods because the trajectory can remain stuck in so-called
superbasins composed of tightly connected sets of states. In this case, many
trajectory segments with the same coordinates are needed before the trajectory
moves on.  Writes only occur for the final state of segments generated by
worker tasks; their magnitude is smaller than reads because the caches ignore
redundant write requests. The number of read and write requests are highest at
the beginning of the run when worker tasks generate segments for the same
state, which is computationally cheap (this motivates
Section~\S\ref{sec:cache-management-using-system-archiecture-knowledge}).

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run (this motivates
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).  The
growth rate, temperature, and number of workers have a predictable effect on
the structure of the keyspace.  Figure~\ref{fig:methodology-keys} shows that
the number of reads changes with different growth rates, but that the spatial
locality of key accesses is similar ({\it e.g.}, some keys are still
\(5\times\) more popular than others).  Figure~\ref{fig:motivation-regimes}
shows how entropy for different growth rates has temporal locality, as the
reads per second for \(\Delta_2\) looks like the reads per second for
\(\Delta_1\) stretched out along the time axis.  Trends also exist for
temperature and number of workers but are omitted here for space. This
structure means that we can learn the regimes and adapt the storage system
(this motivates Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}).

### 2017-11-09 paper/sections/3_methodology.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Mantle Engine: Methodology}

\subsection{Experimental Setup}

\subsection{Mantle: Dynamic Load Balancing Policies}

\begin{itemize}
  \item motivation: Mochi load balancer microservice
  \item background: CephFS implementation
  \item library architecture, callbacks, environment
\end{itemize}

@@

\begin{itemize}
  \item providing environment of metrics
  \item identifying where policies are made
\end{itemize}

>>>>

\section{Methodology}

%  \item background: CephFS implementation
To explore dynamic cache management policies that change during the run, we use
a data management language and policy engine presented
in~\cite{sevilla:sc15-mantle}. The prototype in that paper was called Mantle
and it was originally built on the Ceph file system so administrators could
control file system metadata load balancing policies. We now refer to Mantle as
a policy engine that supports our data management language.  The basic premise
is that load balancing policies can be expressed with a simple API consisting
of ``when", ``where", and ``how much". ``when" controls how aggressive or
conservative the decisions are; ``where" lets the policy control how
distributed or concentrated the data should be; and ``how much" controls the
amount of data that should be sent.

The succinctness of the API lets users inject multiple, dynamic policies. In
this work we focus on a single node, so the ``where" policy is not used. When
we move ParSplice to a distributed key-value store back-end, the ``where" policy
can be used determine which key-value pairs should be moved to which node.

\subsection{Extracting Mantle as a Library}

%  \item motivation: Mochi load balancer microservice
%  \item library architecture, callbacks, environment

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a service ({\it e.g.}, file system, key-value store, application).
Administrators write policies from different domains, like statistics, machine
learning, or storage systems, that that are embedded into the runtime. When the
service makes decisions it executes the administrator-defined policies for
when/where/how much and returns a decision. To do this, the service needs to be
modified to (1) provide an environment of metrics and (2) identify where
policies are set. These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/mantle.png}\\

  \caption{Extracting Mantle as library.\label{fig:mantle}}

\end{figure}

\begin{table}
  \centering
  \begin{tabular}{ r | l | l }
  Metrics     & Data Structure & Description \\\hline
  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
              & resource util. for servers \\
  Time Series & [(ts, val), ..., (ts, val)]
              & accesses by timestamp (ts) \\
  && \\
              & Service      & Example \\\hline
  Cluster     & File Systems & CPU util., Inode reads \\
              & ParSplice    & CPU util., Cache size \\
  Time Series & File Systems & Accesses to directory \\
              & ParSplice    & Accesses to key in DB\\
  \end{tabular}
  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
\end{table}

% policies themselves
\subsubsection{Policies Written as Callbacks} administrators write policies in
Lua that are executed whenever the service needs to make a data management
decision. We chose Lua for simplicity, performance, and portability; it is a
scripting language with simple syntax, which allows administrators to focus on
the policies themselves; it was designed as an embeddable language, so it is
lightweight and does less type checking; and it interfaces nicely with C/C++.
The pink box in Figure~\ref{fig:mantle} shows an example policy for the
``when()" callback in a distributed service, where the current server
(\texttt{whoami}) migrates load if it is has load (\texttt{>0.1}) and if its
neighbor server (\texttt{whoami + 1}) does not have load (\texttt{<0.1}). The
load is calculated using the metrics provided by the environment.

%of timestamp (ts), value pairs representing accesses over time types of
\subsubsection{Environment of Metrics} services can expose cluster metrics for
describing resource utilization and time series metrics for describing accesses
to some data structure over time. Table~\ref{table:metrics} shows how these
metrics are accessed from the policies written by administrators.

For cluster metrics, the service passes a dictionary to Mantle, indexed by
server name and metric. Policies access the cluster metric values by indexing
into a Lua table using \texttt{server} and \texttt{metric}, where server is a
node identifier ({e.g.}, MPI Rank, metadata server name) and metric is a
resource name.  For example, a policy written for an MPI-based service can
access the CPU utilization of the first rank in a communication group using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
load = servers[0]['cpu']
\end{minted}
\end{figure}

Other examples metrics used for file system metadata load balancing, such as
memory pressure, are shown in Figure~\ref{fig:mantle}. The measurements and
exchange of metrics between servers is done by the service; Mantle in CephFS
leverages metrics from other servers collected using CephFS's heartbeats.

For time series metrics, the service passes array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses. For
example, a policy that uses accesses to a certain directory to make migration
decisions collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
ts = array()
for i=1,arraysize() do
  for time,value in string.gmatch(ts, (%w+=%w)) do
    if value == 'mydirectory' then
      count = count + 1
    end
  end
end
\end{minted}
\end{figure}

The service uses a pointer to the time series to facilitate time series with
many values, like accesses to a database or directory in the file system
namespace. This decision limits the time series metrics to only include values
from the {\it current } node, although this is not a limitation of Mantle
itself.

@@

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:parsplice-keyspace-analysis} we only end up using the cache
size although the other metrics proved to be valuable debugging tools. Using
Mantle time series metrics, we expose accesses to the cache as a list of
\texttt{timestamp, keyid} pairs. In
Section~\S\ref{sec:cache-management-using-domain-specific-knowledge}, we
explore a regime detection algorithm that uses this metric.

% where policies are made
We link Mantle directly into the worker cache circled in red in
Figure~\ref{fig:parsplice}. We put the ``when" and ``how much" callbacks
alongside code that checks for memory pressure. It is executed right before the
worker processes incoming and outgoing put/get transactions to the cache. As
stated previously, we do not use the ``where" part of Mantle because we focus
on a single node, but this part of the API will be used when we move the caches
and storage nodes to a key-values store back-end that has load balancing and key
redistribution.

### 2017-11-09 paper/sections/4_dynamic-policies.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on a single in-memory database node using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
implemented as the ``when" and ``how much" policies from the Mantle API; the
``where" callback does not make sense for a single node (segments are either in
the cache or they are not). The evaluation metric is the accuracy and runtime
of each strategy; the strategy should be accurate enough so as to sacrifice
negligible performance and fast tenough to run as often as we want to detect
regimes.

First we sized the cache according to our system specific knowledge. By
``system", we mean the hardware and software of the storage hierarchy. We look
at request rate, unique keys in a sliding window, and bandwidth capabilties. For
example, we know that LevelDB cannot handle high IO request rates.

@@

% results: cache size trade-offs
\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\

  \caption{The performance and resource utilization trade-off for different
cache sizes. ``Baseline" is
ParSplice unmodified and the ``Policy: Constrain Cache Size" graph limits the
size of the cache to save memory.  \label{fig:methodology-tradeoff}}

@@

% what is mantle
To explore dynamic load balancing policies ({\it i.e.} policies that change
during the run), we use the Mantle approach.  Mantle is a framework built on the
Ceph file system that lets administrators control file system metadata load
balancing policies. The basic premise is that load balancing policies can be
expressed with a simple API consisting of ``when", ``where", and ``how much".
The succinctness of the API lets users inject multiple, dynamic policies.

@@

Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using the
Mantle API to program a dynamic load balancing policy into ParSplice that
switches between two policies:

@@

  \caption{Key activity for a 4 hour run shows groups of accesses to the same
  subset of keys. Detecting this ``access regimes" leads to a more accurate cache
  management strategy.\label{fig:keyspace-zoomed}}

>>>>

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on a single in-memory database node using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
implemented as the ``when" and ``how much" policies from the Mantle API; the
``where" callback does not make sense for a single node (segments are either in
the cache or they are not). The evaluation metric is the accuracy and runtime
of each strategy; the strategy should be accurate enough so as to sacrifice
negligible performance and fast enough to run as often as we want to detect
regimes.

First we sized the cache according to our system specific knowledge. By
``system", we mean the hardware and software of the storage hierarchy. We look
at request rate, unique keys in a sliding window, and bandwidth capabilities. For
example, we know that LevelDB cannot handle high IO request rates.

@@

% results: cache size trade-offs
\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
  \caption{The performance and resource utilization trade-off for different
  cache sizes. ``Baseline" is
  ParSplice unmodified and the ``Policy: Constrain Cache Size" graph limits the
  size of the cache to save memory.  \label{fig:methodology-tradeoff}}
\end{figure}

@@

@@

Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
to program a dynamic load balancing policy into ParSplice that switches between
two policies:

@@

  \caption{Key activity for a 4 hour run shows groups of accesses to the same
  subset of keys. Detecting these ``access regimes" leads to a more accurate cache
  management strategy.\label{fig:keyspace-zoomed}}

### 2017-11-09 paper/sections/5_file-systems.tex
paper: integrate Carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Relation to File Systems}

The code snippets in Figures~\ref{lua:parsplice} and~\ref{lua:cephfs} are the
policies used in ParSplice and CephFS, respectively.  ParSplice uses policies
to manage its caches and CephFS uses policies to control load balancing, but
they both can be expressed with the Mantle API.  From a high-level the
ParSplice policy trims the cache if the cache reaches a certain size {\it and}
if it has already absorbed the initial burstiness of the workload; the CephFS
policy migrates load if the metadata load is higher than the average load {\it
and} the current load has been overloaded for more than two iterations.

@@

\textbf{State Persisted Across Decisions} (Fig.~\ref{lua:parsplice}:Lines 5,6;
Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use the Mantle API to
write/read state from previous decisions.  For ParSplice, we save a boolean
that indicates whether we have absorbed the workload's initial burstiness. For
CephFS, we save the number of consecutive instances that the server has been
overloaded. We also clear the count (Line 10) if the server is no longer
overloaded. The underlying implementation saves the values to local disk.

@@

\begin{figure}[tb]
\begin{minted}[linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    if servers[whoami]["load"] > absorb_target then
      WRstate(1)
    end
    if RDstate() == 1 then
      return true
    end
  end
  return false
end
\end{minted}
\caption{ParSplice cache management policy.\label{lua:parsplice}}
\end{figure}

\begin{figure}[tb]
\begin{minted}[linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    overloaded = RDstate() + 1
    WRstate(overloaded)
    if overloaded > 2 then
      return true
    end
  end
  else then
    WRstate(0)
  end
  return false
end
\end{minted}
\caption{CephFS file system metadta load balancer.\label{lua:cephfs}}
\end{figure}

>>>>

\section{Relation to File Systems}
\label{sec:relation-to-file-systems}

The code snippets in Figures~\ref{lua:parsplice} and~\ref{lua:cephfs} are the
policies used in ParSplice and CephFS, respectively.  ParSplice uses policies
to manage its caches and CephFS uses policies to control load balancing, but
they both can be expressed with Mantle.  From a high-level the ParSplice policy
trims the cache if the cache reaches a certain size {\it and} if it has already
absorbed the initial burstiness of the workload; the CephFS policy migrates
load if the metadata load is higher than the average load {\it and} the current
load has been overloaded for more than two iterations.

@@

\textbf{State Persisted Across Decisions} (Fig.~\ref{lua:parsplice}:Lines 5,6;
Fig~\ref{lua:cephfs}:Lines 3,4,10) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded. The
underlying implementation saves the values to local disk.

@@

\begin{figure}[tb]
\footnotesize
\begin{minted}[linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    if servers[whoami]["load"] > absorb_target then
      WRstate(1)
    end
    if RDstate() == 1 then
      return true
    end
  end
  return false
end
\end{minted}
\caption{ParSplice cache management policy.\label{lua:parsplice}}
\end{figure}

\begin{figure}[tb]
\footnotesize
\begin{minted}[linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    overloaded = RDstate() + 1
    WRstate(overloaded)
    if overloaded > 2 then
      return true
    end
  end
  else then
    WRstate(0)
  end
  return false
end
\end{minted}
\caption{CephFS file system metadata load balancer.\label{lua:cephfs}}
\end{figure}

### 2017-11-09 paper/sections/6_conclusion.tex
paper: addresses #7, spell check, intro/methodology hacking
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

>>>>

Data management encompasses a wide range of techniques that vary by domain and
service. Yet, the techniques require policies that shape the decision making
and finding the best policies is a difficult, multi-dimensional problem. We
observe that many of the primitives and resulting strategies have enough in
common that they can be expressed with similar semantics. We present a data
management language and policy engine, called Mantle that is general enough to
express complicated, dynamic policies for two different domains and services.
Rather than attempting to construct a single, complex load balancing policy
that works for a variety of scenarios, we instead use the Mantle framework to
enable software-defined storage systems to flexibly change policies as the
workload changes over time.  In our analysis of the ParSplice key-value
workload we have detected clear workload regimes that are sensitive to the
initial conditions and the scale and duration of the simulation. We have also
demonstrated that changing load balancing policies at runtime in response to
the current workload is an effective mechanism to providing better load
distribution.  Finally, we have demonstrated that Mantle is flexible enough to
support domain-specific knowledge, which lays the groundwork for future work in
adaptable policies and policy generation.
